---
title: "TEP to SSB"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: FALSE
    code_folding: show
    number_sections: TRUE
---

# SET-UP

```{r markdown, echo=F}
library(knitr)
opts_chunk$set(echo = T, collapse = T, fig.align = "center", fig.width = 9, fig.height = 6)
options(width = 300)
```    

```{r settings, message = F,warning=FALSE}
source('0.0_settings.R')
```

# Input data
## Read
```{r read, message = F}
# TEP
repo <- "https://github.com/iml-mackerel/00.0_model/blob/master/"

ys <- c(2020,2022)
tep <- ldply(ys,function(x){
    y <- read.ices(url(paste0(repo,'data/',x,'/tep.dat',"?raw=true")))[[1]]
    data.frame(year=as.numeric(row.names(y)),tep=y[,1],assess=x)
})

# SSB
ys <- c(2016,2018)
ssb <- ldply(ys,function(x){
    y <- read.ices(url(paste0(repo,'data/',x,'/survey.dat',"?raw=true")))[[1]]
    data.frame(year=as.numeric(row.names(y)),ssb=y[,1],assess=x)
})


# bio data
new <- FALSE
if(new){
    bio <- get.bio(species='maquereau',user=imlp.user,password=imlp.pass)
    f <- paste0('Rdata/bio_',Sys.Date(),'.Rdata')
    save(bio,file=f)
}else{
    df <- file.info(list.files("Rdata/", full.names = T,pattern="bio_"))
    f <- rownames(df)[which.max(df$mtime)]
    load(f)
}
print(f)
```

## Subsetting and outlier removal (weight, gonad mass)

```{r outliers, message = F}
# subset
bio.index <- bio[!is.na(bio$weight) & bio$month <8,c('year','month','date','gear','nafo','weight','agef','sample.id','length.frozen','wgonad','matur','sex')]
names(bio.index)[7] <- 'age'
names(bio.index)[9] <- 'length'
nrow(bio.index)

# remove outliers (if length or age do not match weight: remove)
bio.index <- bio.index[bio.index$weight>0.03,]
bio.index <- bio.index[bio.index$age<18,]
nrow(bio.index)

lw <- lm(log(weight)~log(length),data=bio.index[!is.na(bio.index$length),])
bio.index[!is.na(bio.index$length),'outlierLW'] <- abs(residuals(lw))>sd(residuals(lw))*3
bio.index[is.na(bio.index$outlierLW),'outlierLW'] <- FALSE
bio.index <- bio.index[!bio.index$outlierLW,]
nrow(bio.index)

bio.index$gsi <- with(bio.index,wgonad/weight*100)  # hist(bio.index$gsi,1000)
```


# Calculation
Notes:

1.    Used a sex ratio of 0.5 instead of an estimated one.
2.    Average was always taken across all indiduals from a length-stratified subsample...

```{r calcs, message = F}
# average fish weight
w <- ddply(bio.index,c('year'),summarise,w=mean(weight)) # gr

# average fecundity
d <- bio.index[bio.index$matur==5 &!is.na(bio.index$matur) & bio.index$sex=='F',]
d$fec <- 10^(4.32+0.75*log10(d$wgonad*1000))  # gonad weight in gr
fec <- ddply(d,c('year'),summarise,fec=mean(fec,na.rm=TRUE))
fec <- rbind(fec,data.frame(year=2022,fec=mean(tail(fec$fec,5))))  # no gonad mass data in 2022 because of lack of early samples!

# sex ratio
sexr <- 0.5

# get SSB
ssbnew <- merge(merge(tep,w),fec)
ssbnew$ssb <- round(with(ssbnew,tep*10^12*w/(fec*10^3*sexr)),0)
```



# Plots {.tabset}

## TEP
2020: for 1979-2018 based on SSB index divided by W and F (reestimated using old method), for 2019-2020 there are directly new values (still using krigging, etc.).
2022: New time-series (spatial modelling only to fill in gaps, new model to estimate the proportion of eggs spawned).
```{r TEP}
ggplot(tep,aes(x=year,y=tep,col=as.factor(assess)))+geom_point()+geom_line()+labs(col='Assessment',x='Year',y='TEP (billions)')
```

## SSB
2016 and 2018 overlap (because in 2018 only 2017 and 2018 added)
```{r SSB}
ssbs <- rbind(ssb,ssbnew[,names(ssb)])

ggplot(ssbs,aes(x=year,y=ssb,col=as.factor(assess)))+geom_point()+geom_line()+labs(col='Assessment',x='Year',y='SSB')
```

## fec + weight
weight decrease because younger fish.
```{r par}
grid.arrange(
   ggplot(ssbnew,aes(x=year,y=w))+geom_point()+geom_line(),
   ggplot(ssbnew,aes(x=year,y=fec))+geom_point()+geom_line()    
)

```

# Output
```{r csv, message = F}
ssbc <- dcast(ssbs,year~assess,value.var='ssb')
write.csv(ssbc,'csv/ssb.csv',row.names = FALSE)
write.csv(ssbnew,'csv/ssbpara.csv',row.names = FALSE)

kable(ssbc)
```


